Transcriptomic Analysis for Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-07-05 main_redMarked.tex
final main.tex and main_redmared.tex

%\begin{MyColorPar}{red}
The Cancer Genome Atlas (\acrshort{tcga}) profiled \acrshort{hnscc} (528 participants) clinical and genomic data, which were standardized and are available at a unified data portal, \acrfull{gdc} of the \acrfull{nci}.
%\cite{SeanDavis;MartinMorgan2020}
%\end{MyColorPar} % red; the rest moved into materials and methods
% to summarize the advantages of its use.
The highlight of the advantages of applying the TCGA data for cancer biomarker identification includes:
\begin{outline}
\1  To the best of our knowledge, the TCGA database is the largest collection (in both cancer types and cohort size) of comprehensive genomics with survival data available in the field of cancer research so far.  
There are many physical and social features of patients available for survival modeling. % ($X_1 ... X_n$)
The whole-genome sequencing data were harmonized across all \acrfull{gdac}s.  Many databases adopt the essential demographic data from TCGA since it has comprehensive clinical features, such as exposure to alcohol, asbestos, radioactive radon, tobacco smoking, or cigarettes.
\1  TCGA has a remarkable advantage for computational and life scientists who study cancer since useful web-based tools and API are ready to analyze and visualize TCGA data. It might be getting help quickly from the research community for trouble-shooting.
\1  Many achievements of diagnoses, treatments, and prevention with the TCGA data were already published and keep growing\cite{Tomczak2015}.\\[0.1cm]
\end{outline}

@@

They have several genes that are differentially expressed between tumor and normal samples.
It would be helpful to show that alteration in gene expression correlates with phenotypes of HNSCC

@@

Some researchers\cite{Loraine2015a}\cite{Tonella2017a}\cite{Zhao2018}\cite{Li2018a}\cite{Huang2019}\cite{Shen2019}\cite{Schmitt2019}\cite{Xu2021a} tried to find \acrfull{degs} of the HNSCC samples at both genotypic and phenotypic levels (without survival information) for biomarker discovery.
Gene expression data was downloaded from the TCGA or Gene Expression Omnibus (GEO) database (e.g., GSE117973\cite{Schmitt2019}, HIPO-HNC cohort has n = 87).

@@

They used the Database for Annotation, Visualization, and Integrated Discovery (DAVID, available at https://david.ncifcrf.gov/) to obtain information for Gene Ontology (GO), including biological processes, the cellular component, and molecular function. 
The Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway analysis was also used to annotate their potential function.

@@

The protein-protein interaction (PPI) network was conducted by STRING (available at https://string-db.org).
The pathway enrichment analysis of \acrshort{degs} was also performed by DAVID, String, or Cytoscape software programs\cite{Huang2019}\cite{Shen2019}.
Li\cite{Li2018a} and his colleagues made an R package (GDCRNATool) for the implementation of those workflows on gene expression analyses of the TCGA.

@@

Xu and his colleagues\cite{Xu2021a} also identified their biomarkers by analysis of differentially-expressed genes (DEGs).
The gene expression was used to select the preliminary biomarkers.
The significant impact on patients' overall survival was evaluated by Kaplan–Meier survival curves with a log-rank test (\textit{P} value $< 0.01$), and univariate Cox regression.

@@

They validated the candidate genes by using the Gene Expression Profiling Interactive Analysis tool (GEPIA1/2 database using TCGA datasets, developed by Zefang Tang and his colleagues\cite{Tang2017a}\cite{Tang2019}

@@

), GEPIA2021\cite{Li2021} (available at http://gepia2021.cancer-pku.cn/), and the web-tools: Human protein atlas (HPA)\cite{Uhlen2017} (please also see the Discussions section "Validation by Web-based Tools" in this article).

@@

The other investigators should get the rationale or revelation of the genes of interest on a specific cancer type. They should upload those genes manually onto bioinformatics tools, such as SurvExpress (http://bioinformatica.mty.itesm.mx:8080/Biomatec/ \newline
SurvivaX.jsp\cite{Aguirre-Gamboa2013}, which has been lost since Oct/2019 and currently out of funds), and analyze the cohort (e.g., \acrshort{tcga}). After downloading the survival results, they could curate plots and tables carefully.
It is not possible to scan the whole human protein-coding genome in this way.
The web-based tools might set a cutoff at the median, 1/4 quantile, or 3/4 quantile for subsequent analyses. There are several visualization software or R packages which deal with cutoff determination\cite{Abel1984}, %CRITLEVEL
such as Prognoscan\cite{Mizuno2009a}, Cutoff Finder\cite{Budczies2012}, Findcut\cite{Chang2017a}, OptimalCutpoints\cite{Cristina2019},  cutpointr (available at https://github.com/thie1e/cutpointr), and cutoffR (available at https://cran.r-project.org/web/
packages/cutoffR). 
However, non of them could perform jointly predictive of survival times and cutoff selection within subgroups of gene expression.
%  %combine the scanning of the protein-coding genes and cutoff optimization programmatically.

%%% ***
In summary, identifying predictive biomarkers for selecting standard-of-care or advanced systemic therapy\cite{Cristina2019} in \acrshort{hnscc} is crucial.
Our approach describes an in-house workflow implemented in the R script, which runs on the Rstudio server.
Its function includes data retrieving and pre-processing, feature selection, sliding-window cutoff mining, Kaplan-Meier survival analysis, Cox proportional hazard modeling, and biomarker discovery.
The independent HNSCC dataset (GSE65858) was used to validate this strategy.
The workflow, shown in Figure \ref{fig:figure1}, scanned 20,500 human protein-coding genes of the \acrshort{tcga} \acrshort{hnscc} cohort to yield a model with biomarker estimate by gene-expression based survival analysis.

@@

We explain aforementioned situation by examples.
%***NDFIP1  % 給他們一個機會
Reviewing the special case of genes, such as NDFIP1, DKC1, PNMA5, and NPB,
we noticed that NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile (median) cutoff, could achieve a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-adjusted \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233} (please see supplementary figure S1, a "S"- or "W"-shaped acute-bending on the \textit{P} value plot). % it is \ref{fig:pvaluePlot_NDFIP1
%Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in supplementary figure S1).
%After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures, scientific-notation=true]{0.0001074233}.
% the effect size
However, it was excluded as a candidate by small effect size (HR = 1.33 in GSE65858) less than 1.8.
%%%  and HR = 1.42 in GSE2837

The other example is IL19. It has a \textit{P} value plot with acute S-curve bending at the median zone, that lets the FDR-adjusted \textit{P} value has large difference between 50\% quantile cutoff (FDR 0.115) and optimal 48\% quantile cutoff (FDR \num{6.54e-6}). This optimal cutoff method could boost it's statistic significance to pass the correction by both FDR and Bonferroni methods. Even IL19 has competed as a candidate by its effect size (HR = 0.472 in TCGA cohort), but failed by the validation with GSE65858 cohort (HR = 0.630 and FDR-KM \textit{P} = 0.031).

@@

First, although \acrshort{tcga} genomics data were harmonized, 
% three things were found during our investigating and data cleaning:
the pre-processing of TCGA RNA-Seq in our workflow has been done as follows: 
\begin{outline}
\1 HNSCC samples without complete clinical information have been removed;
\1 Null expressed genes in more than 30\% of the HNSCC samples have been excluded; % or 50\%
\1 Updated number of protein-coding genes in the TCGA HNSCC is 20500.
\end{outline}
After investigation of mRNA expression dataset obtained through NCI's Firehose API, we found that expression value of two genes (gene ID: 9906 and gene ID: 728661) was saved together under the entity of gene symbol "SLC35E2".
The expression file of SLC35E2 has almost double in size than those of SLC35E1 or SLC35E3.
% http://gepia2.cancer-pku.cn/#survival is SLC35E2; 證明 TCGA has error
% [line 1549]
%HNSCC.mRNA.Exp.SLC35E2.Fire.Rda
%[line 1654]
%for gene correction (20499 to 20500) from SLC35E2 to 2 genes:
%NCBI Reference Sequences (RefSeq) National Center for Biotechnology Information Reference Sequence (NCBI) RefSeq Release 38 (November 7, 2009) 
According to the Human Gene Database (available at https://www.genecards.org/Search/Keyword?\\queryString=SLC35E2), SLC35E2A (Gene ID: 9906) and SLC35E2B (Gene ID: 728661) should be the correct entities for the TCGA HNSCC dataset.
% https://www.ncbi.nlm.nih.gov/gene/?term=(SLC35E2)+AND+%22Homo+sapiens%22
% keyword: (SLC35E2) AND "Homo sapiens" 
SLC35E2 is the previous symbol of SLC35E2A (reference at https://www.genenames.org/data/gene-symbol-report/\#!/hgnc\_id/HGNC:20863).
Thus, we made reassignment of the expression value of SLC35E2A and SLC35E2B and updated the number of protein-coding genes in this TCGA HNSCC dataset from 20499 to 20500.
% updated on 18-May-2021
%\end{outline}

@@

\1 32.2\% event has "one group" issue in confusion matrix of Chi-square test in Cox regression;
\1 21.05\% error occurs by "one group" issue in log-rank test (survdiff or survdiff.fit function in R package "survival") in Kaplan-Meier estimate;
%1) ok [1] "[ 4 / 20083 ] Generating median-cut KM plot of A1BG"
%Error in survdiff.fit(y, groups, strata.keep, rho) : at A1CF
% error at 16115 gene ""; why? 518 values all are -0.04203 at T
%  There is only 1 group 
%This is because of skewed distribution of the expression value. 
% 67/518 =  13% cases >1, while 450 (87%) cases = -0.30949; so median is -0.30949; 
% all cases belong to >= median => 改為 > median 即可
% 2) % survfit 16114 has data set has no non-missing observations: NA's 113/518 cases
% median = NaN :-); 
% solved by cutoff <- median(gset_OS_mRNA2[, i], na.rm = TRUE)
% and NaN imputation as median
\1 0.78\% has unknown reason (so those 159 genes has been excluded in our workflow).
\end{outline}
These technical problems could not be detected prior to program running.
It might be due to skewed distribution of the expression value or even random error in 20500 RNA-Seq data. % heterogeneity

@@

In line with tumor-agnostic research, we plan to explore common biomarkers crossing TCGA diseases. However, the \acrshort{gdc} provided standardized data frames that could not directly fit our workflow's scope. Before the global genes scanning process, it is necessary to re-format, transpose and, merge the 528 patients' clinical datasets and correlated 20,500 expressions of bio-specimen. This process should be carefully curated to confirm the data integrity within the correct definition\cite{Rendleman2019}. We also plan to upgrade our R script of the cutoff engine to the C++ language and source it in the Rstudio server. The high performance of C++ could speed up the critical steps in this workflow involving heavy computation of matrix data\cite{Woodward2020}. Moreover, it is possible to introduce the Rstudio Shiny app (https://shiny.rstudio.com) as a web-based tool (named "pvalueTex" ) packaged with our workflow in the future.

@@

%*** make Figure 5 holistic: Figure_5_holisticCare.pdf
\begin{figure}[hbt!]
\centering
\includegraphics[width=14cm]{Figure_5_holisticCare.pdf}
\bcaption{The concept of holistic care for \acrshort{hnscc} patients.}
{\\Beyond carcinogenesis - Under the mind-brain-body axis, a stressful environment (giant \textcolor{black}{black} arrow) will trigger emotional reception. The subconscious mind (brain) will release stress hormones and inflammation signals in response to the emotion. The physical body's internal environment (cells) alters epigenetic control in gene regulation and mRNA expression. In a long incubation time, the tissue/cells will be transformed into dysplasia then malignancy (e.g., HNSCC) with helping from known carcinogens.\\
Cancer care - Holistic care should take care of cancer patients' spiritual, emotional, physical, and socioeconomic needs. Physical care will be carried out by medication therapy or surgery. After establishing a therapeutic relationship (TR), the physicians' spiritual properties (empathy, sympathy, and compassion) will engage their patients and recover their self-compassion to gain resilience against the disease through their mind-brain-body axis.
Thus, we suggest electric healthcare records (EHR) include physical, pathological, psychological data, and even more spiritual information. The \acrshort{tcga} should collect "holistic features" as well (\textcolor{green}{green} dashed line).}
\label{fig:figure5}
\end{figure}
\clearpage
%meridian and emotion has mutual reaction
%meridian effect tissue and organ
%(autonomic nerve system, myofascial network, prefrontal cortex)
%placebo-effect influence on drug discovery
%gene expression in deep learning
%Ching, T. et al. Opportunities and obstacles for deep learning in biology and medicine. J. R. Soc. Interface 15, (2018).
%慈悲心 利他心
%博士班學生，越研究越發現，科學的不足，心理與靈性的力量，更是背後推動的冥冥力量，身心靈是合一的
%#正念 除了可以減壓，還可以治癒疼痛
% modify the dark environment 改造患者的心，就能改造他的身體環境，有利消除癌症

@@

GDC is available at https://portal.gdc.cancer.gov/projects/TCGA-HNSC.
\acrshort{tcga} offers several computational tools to the public for facilitating cancer research.
Broad \acrfull{gdac} Firebrowse (firebrowse.org, version 1.1.35, 2016\_09\_27) is one of those tools to provide data access to each \acrshort{tcga} disease through a \acrfull{rest} \acrfull{api}.
The 528 \acrshort{tcga} \acrshort{hnscc} patients' clinical information and \acrshort{rnaseq} data were obtained from the Firebrowse \acrshort{rest}ful \acrshort{api} with an R package, FirebrowseR (available at https://github.com/mariodeng/\linebreak
FirebrowseR)\cite{Deng2017}. 
We utilized FirebrowseR with our analysis workflow (see Figure \ref{fig:figure1}, \textcolor{green}{green} square) to receive standardized data frames while avoiding data re-formatting, often causing some errors.

@@

The number of protein-coding genes was suggested as 20,500\cite{Clamp2007}. The \acrshort{gdc} Data Portal provided \acrshort{tcga} data has been harmonized and re-aligned \acrlong{rnaseq} data against an official reference genome build (\acrlong{grch38}, \acrshort{grch38}). \acrshort{rnaseq} expression level read counts produced by Illumina HiSeq have been normalized using the \acrfull{fpkm} method, as described in reference\cite{FPKM2017}.
The \acrshort{rnaseq} preprocessor of Broad \acrshort{gdac} picked the \acrfull{rsem} value from Illumina HiSeq/GA2 messenger \acrshort{rnaseq} level\_3 (v2) dataset of \acrshort{nci} \acrshort{gdc}. It made the messenger \acrshort{rnaseq} matrix with log2 transformed for the downstream analysis, as described in their reference\cite{RSEM2016}.
We utilized FirebrowseR's function call, Samples.mRNASeq(cohort = "HNSC", gene=GeneName, format="csv"), to download each \acrshort{rnaseq} data of all \acrshort{hnscc} patients and to save as 20,499 data frame files, named as "HNSCC.mRNA.Exp.[GeneName].Fire.Rda".
After careful investigation of the genomics dataset, the \acrshort{rnaseq} values of "\acrfull{slca}" and "\acrfull{slcb}" should be considered two distinct expression entities. We concluded that the number of protein-coding genes in the \acrshort{tcga} dataset is 20,500. We removed null expressed genes, which over 50\% of the cohort, to avoid the useless result.

@@

\begin{MyColorPar}{red}
GSE65858 is a dataset for candidate selection in our workflow.
The Gene Expression Omnibus (GEO) database\cite{Chung2006},, founded by \acrfull{ncbi}, is a public repository supporting MIAME-compliant data,  including microarray and sequence-based experiments.
R package of GEOquery\cite{Sean2007} was used to
download RNA-Seq dataset (in SOFT or MINiML format) of a HNSCC cohort, GSE65858, from the GEO database (available at https://www.ncbi.nlm.nih.gov/geo/geo2r/?\\acc=GSE65858). 
GSE65858\cite{Wichmann2015} has OS event, RFS event, and survival time.
There are 270 HNSCC participants involved in this cohort.
%The number of genes (or probes of RNA-Seq in platform GPL10558, Illumina HumanHT-12 V4.0 expression beadchip) is 30,330.
%n=270 n= 273 = ?? + 168(alive) + 88(dead) = 256
The expression data was generated using the platform GPL10558 (Illumina HumanHT-12 v4.0 Expression BeadChip), which targets on more than 30,330 annotated genes (47,000 probes, derived from the \acrshort{ncbi} Reference Sequence, release 38 on 7 November, 2009). % (RefSeq) 
GSE65858 is a dataset for helping of candidate selection in our workflow.
We have also performed Kaplan-Meier (with FDR-correction of \textit{P} value) and Cox survival analyses under the cutoff of each gene expression at their median value.
\end{MyColorPar} % red

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%% optional: The data presented in this study are available in supplementary material. Supplementary_figures.tex
\supplementary{The following are available online at \linksupplementary{Supplementary_figures.pdf}, 
	Figure S1: \textit{P} value plot of the gene NDFIP1,
%	Figure S2: GSE2837 analysis by PrognoScan,  
%	Table S1: The GSE2837 query results from PrognoScan,
	Table S1: 10 bad guys, % ? don't keep it
	Table S2: 10 good guys, % 要加入 CALML5
	Video S1: title. A supporting video article is available at doi: link}

>>>>

%\begin{MyColorPar}{red}
The Cancer Genome Atlas (\acrshort{tcga}) profiled \acrshort{hnscc} (528 participants) clinical and genomic data, which were standardized and are available at a unified data portal, \acrfull{gdc} of the \acrfull{nci}.
%\cite{SeanDavis;MartinMorgan2020}
%\end{MyColorPar} % red; the rest moved into materials and methods
% to summarize the advantages of its use.
The highlight of the advantages of applying the TCGA data for cancer biomarker identification includes:
\begin{outline}
\1  To the best of our knowledge, the TCGA database is the largest collection (in both cancer types and cohort size, especially in HNSCC) of comprehensive genomics with survival data available in the field of cancer research so far.
%There are many physical and social features of patients available for survival modeling. % ($X_1 ... X_n$)
The whole-genome sequencing data were harmonized across all \acrfull{gdac}s.  Many databases adopt the essential demographic data from TCGA since it has comprehensive physical and social features of patients, such as exposure to alcohol, asbestos, radioactive radon, tobacco smoking, or cigarettes.
\1  TCGA has a remarkable advantage for computational and life scientists who study cancer since useful web-based tools and \acrfull{api} are ready to analyze and visualize TCGA data. It might be getting help quickly from the research community for trouble-shooting.
\1  Many achievements of diagnoses, treatments, and prevention from the TCGA data were already published and keep growing\cite{Tomczak2015}.\\[0.1cm]
\end{outline}

@@

%They have several genes that are differentially expressed---\acrfull{degs} between tumor and normal samples.
It would be helpful to show that alteration in gene expression correlates with phenotypes of HNSCC.

@@

Some researchers\cite{Loraine2015a}\cite{Tonella2017a}\cite{Zhao2018}\cite{Li2018a}\\\cite{Huang2019}\cite{Shen2019}\cite{Schmitt2019} tried to find \acrfull{degs} of the HNSCC samples at both genotypic and phenotypic levels (without survival information) for biomarker discovery.
Gene expression data were downloaded from the TCGA or Gene Expression Omnibus (GEO) databases (e.g., GSE117973\cite{Schmitt2019}, HIPO-HNC cohort has n = 87).

@@

They used the Database for Annotation, Visualization, and Integrated Discovery---DAVID (available at https://david.ncifcrf.gov/) to obtain information for Gene Ontology (GO), including biological processes, the cellular component, and molecular function. 
The Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway analysis was also used to annotate the potential functions of their biomarker candidates.

@@

%The protein-protein interaction (PPI) network was conducted by STRING (available at https://string-db.org).
The pathway enrichment analysis of \acrshort{degs} was also performed by DAVID, STRING (available at https://string-db.org), or Cytoscape software\cite{Huang2019}\cite{Shen2019}.
Li\cite{Li2018a} and his colleagues made an R package (GDCRNATool) for the implementation of those workflows for gene expression analyses of the TCGA.

@@

Xu and his colleagues\cite{Xu2021a} also identified their biomarkers by  \acrshort{degs} analysis.
%The differential expression of gene was used to select the preliminary biomarkers.
The significant impact of genes on patients' overall survival was evaluated by Kaplan–Meier survival curves with a log-rank test (\textit{P} value $< 0.01$), and univariate Cox regression.

@@

They validated the candidate genes by using the web-based Tools of Gene Expression Profiling Interactive Analysis tool---GEPIA and Human protein atlas (HPA) databases.
GEPIA was developed, using TCGA datasets, by Zefang Tang and his colleagues (version 1\cite{Tang2017a}, version 2\cite{Tang2019}, and GEPIA2021\cite{Li2021}, available at http://gepia2021.cancer-pku.cn/).

@@

HPA\cite{Uhlen2017} applied \acrfull{ihc} for the TCGA database (please see details in the Discussions section "Validation by Web-based Tools").

@@

%http://bioinformatica.mty.itesm.mx:8080/Biomatec/SurvivaX.jsp
% bioinformatics
The other investigators should get the rationale or revelation of the genes of interest on a specific cancer type. They should upload those genes manually onto web-based tools, such as SurvExpress\cite{Aguirre-Gamboa2013} (available at http://bioinformatica.mty.itesm.mx:8080/\\Biomatec/SurvivaX.jsp), then analyze the cohort of interest (e.g., \acrshort{tcga}). After downloading the survival results, they could curate plots and tables carefully.
It is not possible to scan the whole human protein-coding genome in this way.
The web-based tools might set a cutoff at the median, 1/4 quantile, or 3/4 quantile for subsequent analyses. There are several visualization software or R packages which deal with cutoff determination\cite{Abel1984}, %CRITLEVEL
such as Prognoscan\cite{Mizuno2009a}, Cutoff Finder\cite{Budczies2012}, Findcut\cite{Chang2017a}, OptimalCutpoints\cite{Cristina2019},  cutpointr (available at https://github.com/thie1e/cutpointr), and cutoffR (available at https://cran.r-project.org/web/packages/cutoffR). 
However, non of them could perform jointly survival analysis with cutoff selection and whole-genome scanning.
%  %combine the scanning of the protein-coding genes and cutoff optimization programmatically.

%%% ***
In summary, identifying predictive biomarkers for selecting standard-of-care or advanced systemic therapy\cite{Cristina2019} in \acrshort{hnscc} is crucial.
Our approach describes an in-house workflow implemented in the R script, which runs on the Rstudio server.
Its function includes data retrieving and pre-processing, feature selection, sliding-window cutoff selection, Kaplan-Meier survival analysis, Cox proportional hazard modeling, and biomarker discovery.
The independent HNSCC dataset (GSE65858)\cite{Wichmann2015} was used to validate this strategy.
The workflow, shown in Figure \ref{fig:figure1}, has scanned 20,500 human protein-coding genes of the \acrshort{tcga} \acrshort{hnscc} cohort to yield a model with biomarker estimate by gene-expression based survival analysis.

@@

We explain aforementioned situation by examples.
%***NDFIP1  % 給他們一個機會
Reviewing the special case of genes, such as NDFIP1, DKC1, PNMA5, and NPB,
we noticed that NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile (median) cutoff, could achieve a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-adjusted \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233} (please see supplementary Figure S1, a "S"- or "W"-shaped acute-bending on the \textit{P} value plot). % it is \ref{fig:pvaluePlot_NDFIP1
%Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in supplementary figure S1).
%After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures, scientific-notation=true]{0.0001074233}.
% the effect size
However, it was excluded as a candidate by small effect size (HR = 1.33 in GSE65858) less than 1.8.
%%%  and HR = 1.42 in GSE2837

The other example is IL19. It has a \textit{P} value plot with acute S-curve bending at the median zone, that lets the FDR-adjusted \textit{P} value has large difference between 50\% quantile cutoff (FDR 0.115) and optimal 48\% quantile cutoff (FDR \num{6.54e-6}). This optimal cutoff method could boost it's statistic significance to pass the correction by both FDR and Bonferroni methods. Even IL19 has competed as a candidate by its effect size (HR = 0.472 in TCGA cohort), but failed by the validation with GSE65858 cohort (small effect size as HR = 0.630, and FDR-KM \textit{P} = 0.031).

@@

First, since \acrshort{tcga} genomic data were harmonized, 
% three things were found during our investigating and data cleaning:
the pre-processing of TCGA RNA-Seq in our workflow has been done as follows: 
\begin{outline}
\1 HNSCC samples without complete clinical information have been removed;
\1 Null expressed genes in more than 30\% of the HNSCC samples have been excluded; % or 50\%
\1 Updated number of protein-coding genes in the TCGA HNSCC is 20500.
\end{outline}
After investigation of mRNA expression dataset obtained through NCI's Firehose API, we found that expression value of two genes (gene ID: 9906 and gene ID: 728661) was saved together under the entity of gene symbol "SLC35E2".
The expression file of SLC35E2 has almost double in size than those of SLC35E1 or SLC35E3.
% http://gepia2.cancer-pku.cn/#survival is SLC35E2; 證明 TCGA has error
% [line 1549]
%HNSCC.mRNA.Exp.SLC35E2.Fire.Rda
%[line 1654]
%for gene correction (20499 to 20500) from SLC35E2 to 2 genes:
%NCBI Reference Sequences (RefSeq) National Center for Biotechnology Information Reference Sequence (NCBI) RefSeq Release 38 (November 7, 2009) 
According to the Human Gene Database (available at https://www.genecards.org/Search/Keyword?\\queryString=SLC35E2), SLC35E2A (Gene ID: 9906) and SLC35E2B (Gene ID: 728661) should be the correct entities for the TCGA HNSCC dataset.
% https://www.ncbi.nlm.nih.gov/gene/?term=(SLC35E2)+AND+%22Homo+sapiens%22
% keyword: (SLC35E2) AND "Homo sapiens" 
SLC35E2 is the previous symbol of SLC35E2A (reference at https://www.genenames.org/data/gene-symbol-report/\#!/hgnc\_id/HGNC:20863).
Thus, we made reassignment of the expression value of SLC35E2A and SLC35E2B and updated the number of protein-coding genes in this TCGA HNSCC dataset from 20499 to 20500.
% updated on 18-May-2021
%\end{outline}

@@

\1 32.2\% event has "one group" issue in confusion matrix of Chi-square test in Cox regression;
\1 21.05\% error occurs by "one group" issue in log-rank test (survdiff or survdiff.fit function in R package "survival") in Kaplan-Meier estimate;
%1) ok [1] "[ 4 / 20083 ] Generating median-cut KM plot of A1BG"
%Error in survdiff.fit(y, groups, strata.keep, rho) : at A1CF
% error at 16115 gene ""; why? 518 values all are -0.04203 at T
%  There is only 1 group 
%This is because of skewed distribution of the expression value. 
% 67/518 =  13% cases >1, while 450 (87%) cases = -0.30949; so median is -0.30949; 
% all cases belong to >= median => 改為 > median 即可
% 2) % survfit 16114 has data set has no non-missing observations: NA's 113/518 cases
% median = NaN :-); 
% solved by cutoff <- median(gset_OS_mRNA2[, i], na.rm = TRUE)
% and NaN imputation as median
\1 0.78\% has unknown reason (so those 159 genes has been excluded in our workflow).
\end{outline}
These technical problems could not be detected prior to program running.
It might be due to skewed distribution of the expression value or even random error derived during the RNA sequencing procedure. % heterogeneity

@@

In line with tumor-agnostic research, we plan to explore common biomarkers crossing TCGA diseases. However, the \acrshort{gdc} provided standardized data frames that could not directly fit our workflow's scope. Before the global genes scanning process, it is necessary to re-format, transpose and, merge the 528 patients' clinical datasets and correlated 20,500 expressions of bio-specimen. This process should be carefully curated to confirm the data integrity within the correct definition\cite{Rendleman2019}. We also plan to upgrade our R script of the cutoff engine to the C++ language and source it in the Rstudio server. The high performance of C++ could speed up the critical steps in this workflow involving heavy computation of matrix data\cite{Woodward2020}. Moreover, it is possible to introduce the Rstudio Shiny app (https://shiny.rstudio.com) as a web-based tools (named "pvalueTex" ) packaged with our workflow in the future.

@@

%*** make Figure 5 holistic: Figure_5_holisticCare.pdf
\begin{figure}[hbt!]
\centering
\includegraphics[width=14cm]{Figure_5_holisticCare.pdf}
\bcaption{The concept of holistic care for \acrshort{hnscc} patients.}
{\\Beyond carcinogenesis - Under the mind-brain-body axis, a stressful environment (giant \textcolor{black}{black} arrow) will trigger emotional reception. The subconscious mind (brain) will release stress hormones and inflammation signals in response to the emotion. The physical body's internal environment (cells) alters epigenetic control in gene regulation and mRNA expression. In a long incubation time, the tissue/cells will be transformed into dysplasia then malignancy (e.g., HNSCC) with helping from known carcinogens.\\
Cancer care - Holistic care should take care of cancer patients' spiritual, emotional, physical, and socioeconomic needs. Physical care will be carried out by medication therapy or surgery. After establishing a therapeutic relationship (TR), the physicians' spiritual properties (empathy, sympathy, and compassion) will engage their patients and recover their self-compassion to gain resilience against the disease through their mind-brain-body axis.
Thus, we suggest that electric healthcare records (EHR) should include physical, pathological, psychological data, and even more spiritual information. The \acrshort{tcga} might collect those "holistic features"  (\textcolor{green}{green} dashed line) for further study of personalized medicine.}
\label{fig:figure5}
\end{figure}
\clearpage
%meridian and emotion has mutual reaction
%meridian effect tissue and organ
%(autonomic nerve system, myofascial network, prefrontal cortex)
%placebo-effect influence on drug discovery
%gene expression in deep learning
%Ching, T. et al. Opportunities and obstacles for deep learning in biology and medicine. J. R. Soc. Interface 15, (2018).
%慈悲心 利他心
%博士班學生，越研究越發現，科學的不足，心理與靈性的力量，更是背後推動的冥冥力量，身心靈是合一的
%#正念 除了可以減壓，還可以治癒疼痛
% modify the dark environment 改造患者的心，就能改造他的身體環境，有利消除癌症

@@

GDC is available at https://portal.gdc.cancer.gov/projects/TCGA-HNSC.
\acrshort{tcga} offers several computational tools to the public for facilitating cancer research.
Broad \acrfull{gdac} Firebrowse (firebrowse.org, version 1.1.35, 2016\_09\_27) is one of those tools to provide data access to each \acrshort{tcga} disease through a \acrfull{rest} \acrfull{api}.
The 528 \acrshort{tcga} \acrshort{hnscc} patients' clinical information and \acrshort{rnaseq} data were obtained from the Firebrowse \acrshort{rest}ful \acrshort{api} with an R package, FirebrowseR (available at https://github.com/mariodeng/FirebrowseR)\cite{Deng2017}. 
We utilized FirebrowseR with our analysis workflow (see Figure \ref{fig:figure1}, \textcolor{green}{green} square) to receive standardized data frames while avoiding data re-formatting, often causing some errors.

@@

The number of protein-coding genes was suggested as 20,500\cite{Clamp2007}. The \acrshort{gdc} Data Portal provided \acrshort{tcga} data has been harmonized and re-aligned \acrlong{rnaseq} data against an official reference genome build (\acrlong{grch38}, \acrshort{grch38}). \acrshort{rnaseq} expression level read counts produced by Illumina HiSeq have been normalized using the \acrfull{fpkm} method, as described in reference\cite{FPKM2017}.
The \acrshort{rnaseq} preprocessor of Broad \acrshort{gdac} picked the \acrfull{rsem} value from Illumina HiSeq/GA2 messenger \acrshort{rnaseq} level\_3 (v2) dataset of \acrshort{nci} \acrshort{gdc}. It made the messenger \acrshort{rnaseq} matrix with log2 transformed for the downstream analysis, as described in their reference\cite{RSEM2016}.
We utilized FirebrowseR's function call, Samples.mRNASeq(cohort = "HNSC", gene=GeneName, format="csv"), to download each \acrshort{rnaseq} data of all \acrshort{hnscc} patients and to save as 20,499 data frame files, named as "HNSCC.mRNA.Exp.[GeneName].Fire.Rda".
After careful investigation of the genomics dataset, the \acrshort{rnaseq} values of "\acrfull{slca}" and "\acrfull{slcb}" should be considered as two distinct expression entities. We concluded that the number of protein-coding genes in the \acrshort{tcga} dataset is 20,500. We removed null expressed genes, which over 30\% of the cohort, to avoid the useless result.

@@

\begin{MyColorPar}{red}
GSE65858\cite{Wichmann2015} is a dataset for helping of candidate selection in our workflow.
The Gene Expression Omnibus (GEO) database\cite{Chung2006}, founded by \acrfull{ncbi}, is a public repository supporting MIAME-compliant data,  including microarray and sequence-based experiments.
R package of GEOquery\cite{Sean2007} was used to
download RNA-Seq dataset (in SOFT or MINiML format) of a HNSCC cohort, GSE65858, from the GEO database (available at https://www.ncbi.nlm.nih.gov/geo/\\geo2r/?acc=GSE65858). 
GSE65858 has OS event, RFS event, and survival time.
There are 270 HNSCC participants involved in this cohort.
%The number of genes (or probes of RNA-Seq in platform GPL10558, Illumina HumanHT-12 V4.0 expression beadchip) is 30,330.
%n=270 n= 273 = ?? + 168(alive) + 88(dead) = 256
The expression data was generated using the platform GPL10558 (Illumina HumanHT-12 v4.0 Expression BeadChip), which targets on more than 30,330 annotated genes (47,000 probes, derived from the \acrshort{ncbi} Reference Sequence, release 38 on 7 November, 2009). % (RefSeq) 
We have performed Kaplan-Meier (with FDR-correction of \textit{P} value) and Cox survival analyses under the cutoff of each gene expression at their median value. The biomarker candidates were a consensus result of TCGA and GSE65858 analyses.
\end{MyColorPar} % red

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%% optional: The data presented in this study are available in supplementary material. Supplementary_figures.tex
\supplementary{The following are available online at \linksupplementary{Supplementary_figures.pdf}\\
	Figure S1: \textit{P} value plot of the gene NDFIP1\\
%	Figure S2: GSE2837 analysis by PrognoScan,  
%	Table S1: The GSE2837 query results from PrognoScan,
	Table S1: 10 genes have unfavorable prognosis\\ % ? don't keep it
	Table S2: 10 genes have favorable prognosis\\} % 要加入 CALML5
%	Video S1: title. A supporting video article is available at doi: link}
